---
title: "Introduction to SABER MCMC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SABER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(SABER)
library(plotly)
library(BayesianTools)
library(stringr)
library(dplyr)
library(tidyr)
library(lubridate)
```

## Overview

SABER is an R package for Bayesian inversion of radiative transfer models using in-situ data and prior distribution fitting.

## Fully unconstrained inversion with uniform priors

```{r}
rrs <- rrs_hw %>%
  filter(ensemble == 1657138120) %>%
  mutate(rrs_0m = rrs_0p_to_0m(rrs_estimate)) %>%
  select(wavelength, rrs_0m)
```

```{r}
# How to retrieve bottom reflectance without specifying a class ?
# Is it feasible ? Definitely seems useful.
par_fixed <- c(
  "water_type" = 2,
  "theta_view" = 0,
  "theta_sun" = 20,
  "r_rs_b_saccharina" = 1,
  "sd" = 1
)

par_estimates <- inverse_mcmc(
  rrs = rrs,
  forward_model = "am03",
  par_inversed = c("chl", "a_g_440", "bb_p_550", "h_w"),
  prior = NULL,
  lower = c(0, 0, 0, 0),
  upper = c(100, 10, 0.5, 10),
  par_fixed = par_fixed,
  iterations = 10000,
  burnin = 2000,
  sampler = "DEzs"
)

# plot(out)
```

Now we can compare what those OAC gives in "am03" and compare that with our input rrs.

```{r}
# par_estimates <- tibble(
#   name = names(MAP(out)[[1]]),
#   value = MAP(out)[[1]]
# ) %>%
#   filter(name %in% c("chl", "a_g_440", "bb_p_550", "h_w"))
#
# par_estimates <- stats::setNames(par_estimates$value, par_estimates$name)

iop <- iop_from_oac(
  rrs$wavelength,
  par_estimates
)


r_b_fraction_vec <- par_fixed[grep("^r_rs_b", names(par_fixed))]

rrs_bottom <- compute_r_rs_b_lmm(
  fractions = r_b_fraction_vec
)

rrs_retrieved <- forward_am03(
  # wavelength = rrs$wavelength,
  # iop = iop,
  # water_type = 2,
  # theta_view = 0,
  # theta_sun = 20,
  # optically_shallow = T,
  # h_w = 2,
  # rrs_bottom = rrs_bottom,
  # verbose = T
  wavelength = rrs$wavelength,
  iop = iop,
  water_type = 2,
  theta_view = 0,
  theta_sun = 20,
  h_w = par_estimates["h_w"],
  r_b = rrs_bottom
)

ply <- plot_ly() %>%
  add_lines(x = rrs$wavelength, y = rrs$rrs_0m, name = "rrs_observed") %>%
  add_lines(x = rrs$wavelength, y = rrs_retrieved, name = "rrs_retrieved")

ply
```
### Batch processing

```{r}
ensembles <- unique(rrs_hw$ensemble)

ensembles <- ensembles[
  ensembles > as.numeric(ymd_hms("2022-07-06 20:58:20")) &
    ensembles > as.numeric(ymd_hms("2022-07-06 21:04:50"))
]

rrs_hw_sub <- rrs_hw %>%
  # filter(ensemble %in% ensembles) %>%
  filter(wavelength < 725) %>%
  mutate(
    rrs_0m = rrs_0p_to_0m(rrs_estimate),
    h_w = altitude_m + abs(bottom_elevation_m)
  )

ply <- plot_ly(rrs_hw_sub) %>%
  add_lines(x = ~wavelength, y = ~rrs_0m, name = ~ paste0(ensemble, "_", round(plant_height_m, 2)))

ply
```

```{r}
par_fixed <- c(
  "water_type" = 2,
  "theta_view" = 0,
  "theta_sun" = 20,
  "sd" = 1
)

inversed_data <- rrs_hw_sub %>%
  group_by(ensemble) %>%
  nest() %>%
  mutate(
    par_inversed = purrr::map(
      .x = data,
      ~ inverse_mcmc(
        rrs = .x,
        forward_model = "am03",
        par_inversed = c("chl", "a_g_440", "bb_p_550", "h_w", "r_rs_b_saccharina", "r_rs_b_cca", "r_rs_b_gravel"),
        prior = NULL,
        lower = c(0, 0, 0, 0, 0, 0, 0),
        upper = c(100, 10, 0.5, 10, 1, 1, 1),
        par_fixed = par_fixed,
        iterations = 10000,
        burnin = 2000,
        sampler = "DEzs"
      )
    )
  )
```

```{r}
par_inversed <- inversed_data$par_inversed %>%
  bind_rows() %>%
  rename_with(.fn = ~ paste0(.x, "_estim")) %>%
  mutate(ensemble = inversed_data$ensemble, .before = 1)

compare_h_w <- rrs_hw_sub %>%
  select(ensemble, h_w, plant_height_m, percent_coverage) %>%
  unique() %>%
  inner_join(par_inversed, by = c("ensemble"))
```

```{r}
ply <- plot_ly(compare_h_w) %>%
  add_markers(x = ~h_w, y = ~h_w_estim, color = ~plant_height_m) %>%
  layout(shapes = list(list(
    type = "line",
    x0 = ~ min(h_w, h_w_estim),
    x1 = ~ max(h_w, h_w_estim),
    xref = "x",
    y0 = ~ min(h_w, h_w_estim),
    y1 = ~ max(h_w, h_w_estim),
    yref = "y",
    line = list(color = "black")
  )))

ply
```

## Load Sample Data

```{r}
# data("nomad_sample")
# head(nomad_sample)
nomad_sample <- sample_nomad()
```

## Fit Prior Distributions

```{r}
chl_prior <- make_prior(nomad_sample$chl, dist = "weibull", name = "chl")
ag_prior <- make_prior(nomad_sample$ag443, dist = "weibull", name = "ag_440")
# ad_prior <- make_prior(nomad_sample$ad443, dist = "weibull", name = "ad440")
bbp_prior <- make_prior(nomad_sample$bbp555, dist = "weibull", name = "bbp_550")

priors <- list(chl_prior, ag_prior, bbp_prior)
```

## Load Rrs

```{r}
rrs <- readr::read_csv(fs::path_package("SABER", "data", "demo_rrs_Om.csv"))

par_fixed <- tibble(
  parameter = c("water_type", "theta_view", "theta_sun", "optically_shallow", "h_w", "rb_class_1", "rb_class_2", "rb_class_3", "sd"),
  value = c(2, 0, 20, T, 2, 0.25, 0.25, 0.5, 0.05),
)
```

## Bundle Priors and Define Likelihood

```{r}
prior_bundle <- make_prior_bundle(
  priors = priors,
  lower = c(0, 0, 0),
  upper = c(30, 5, 0.5)
)

likelihood <- make_likelihood(
  model = "am03",
  error = "log-ll",
  rrs_observed = rrs,
  par_fixed = par_fixed
)
```

## Run Inversion

```{r, eval=FALSE}
out <- inverse_mcmc(
  prior_bundle,
  likelihood,
  iterations = 10000,
  burnin = 2000,
  sampler = "DEzs",
  use_prior = F
)

summary(out)
plot(out, start = 2000)
```

```{r}
oac_estimates <- tibble(
  name = names(MAP(out)[[1]]),
  value = MAP(out)[[1]]
) %>%
  filter(name %in% c("chl", "ag_440", "bbp_550")) %>%
  tidyr::pivot_wider(names_from = "name", values_from = "value")


iop <- iop_from_oac(
  rrs$wavelength,
  oac_estimates,
  verbose = T
)

end_member_fraction <- par_fixed %>%
  filter(str_detect(parameter, "rb_"))

rrs_bottom <- rrs_bottom_lmm(
  rrs_end_member = rb_algae_wise,
  end_member_fraction = end_member_fraction$value,
  wavelength = rrs$wavelength,
  verbose = FALSE
)

rrs_retrieved <- forward_am03(
  wavelength = rrs$wavelength,
  iop = iop,
  water_type = 2,
  theta_view = 0,
  theta_sun = 20,
  optically_shallow = T,
  h_w = 2,
  rrs_bottom = rrs_bottom,
  verbose = T
)

ply <- plot_ly() %>%
  add_lines(x = rrs$wavelength, y = rrs$rrs_0m, name = "rrs_observed") %>%
  add_lines(x = rrs_retrieved$wavelength, y = rrs_retrieved$rrs_0m, name = "rrs_retrieved")


ply
```

