---
title: "Introduction to SABER MCMC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SABER}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# library(SABER)
devtools::load_all(".")
library(plotly)
library(BayesianTools)
library(stringr)
library(dplyr)
library(tidyr)
library(lubridate)
library(cmdstanr)

font <- list(family = "Times New Roman", size = 20, weight = "normal")

border <- list(
  type = "rect",
  fillcolor = "transparent",
  line = list(width = 0.5),
  xref = "paper",
  yref = "paper",
  x0 = 0,
  x1 = 1,
  y0 = 0,
  y1 = 1
)

library(reticulate)
use_condaenv("kaleido", required = TRUE)
py_config()
py_run_string("import sys")

here <- function(...) {
  start <- "/D/Documents/phd/thesis"
  file.path(start, ...)
}
```

## Overview

SABER is an R package for Bayesian inversion of radiative transfer models using in-situ data and prior distribution fitting.

### Compile STAN model

```{r}
# model <- cmdstan_model("/home/raphael/CLionProjects/saber-lib/src/model.stan")

model <- cmdstan_model(
  stan_file = "/home/raphael/CLionProjects/saber-lib/src/model.stan",
  user_header = "/home/raphael/CLionProjects/saber-lib/src/rtm_stan_funcs.hpp"
)
```

## Fully unconstrained inversion with uniform priors

```{r}
rrs <- rrs_hw %>%
  filter(
    wavelength > 400, wavelength < 720,
    # ensemble == 1657138160
    ensemble == 1657138460
    ) %>%
  mutate(rrs_0m = rrs_0p_to_0m(rrs_estimate)) %>%
  select(wavelength, rrs_0m, rrs_unc)
```


```{r}
library(dplyr)
library(tidyr)

interp1 <- function(x, y, xout) {
  approx(x = x, y = y, xout = xout, rule = 2, ties = "ordered")$y
}

# Main builder -------------------------------------------------------------
build_saber_stan_luts <- function(wavelength,
                                 a_w_df,
                                 a0_a1_df,
                                 r_rs_b_long,
                                 bb_w_fun = function(wl) 0.00111 * (wl / 500)^(-4.32)) {

  stopifnot(is.numeric(wavelength), length(wavelength) >= 2)

  # ---- a_w ----
  a_w <- interp1(a_w_df$wavelength, a_w_df$a_w, wavelength)

  # ---- a0/a1 ----
  a0 <- interp1(a0_a1_df$wavelength, a0_a1_df$a0, wavelength)
  a1 <- interp1(a0_a1_df$wavelength, a0_a1_df$a1, wavelength)

  # ---- bb_w (pure water backscatter) ----
  bb_w <- bb_w_fun(wavelength)

  # ---- bottom library: long -> wide matrix ----
  # Expect columns: class, wavelength, r_rs_b_mean (adjust names below if needed)
  rb <- r_rs_b_long %>%
    transmute(class = as.character(.data$class),
              wl    = .data$wavelength,
              mean  = .data$r_rs_b_mean)

  classes <- rb %>%
    distinct(class) %>%
    arrange(class) %>%
    pull(class)

  rb_wide <- rb %>%
    pivot_wider(names_from = class, values_from = mean) %>%
    arrange(wl)

  rb_wl_master <- rb_wide$wl

  r_rs_b_lib <- sapply(classes, function(cl) {
    interp1(rb_wl_master, rb_wide[[cl]], wavelength)
  })

  r_rs_b_lib <- matrix(r_rs_b_lib, nrow = length(wavelength), ncol = length(classes))
  colnames(r_rs_b_lib) <- classes

  list(
    a_w = a_w,
    a0 = a0,
    a1 = a1,
    bb_w = bb_w,
    r_rs_b_lib = r_rs_b_lib,
    classes = classes
  )
}

make_saber_stan_data <- function(rrs_obs, wavelength, sigma,
                                water_type, theta_sun_deg, theta_view_deg, shallow,
                                bottom_class_names,
                                pkgname = "SABER") {

  # load datasets from package (like your .onLoad did)
  a_w_df     <- get(data("a_w", package = pkgname), envir = environment())
  a0_a1_df   <- get(data("a0_a1_phyto", package = pkgname), envir = environment())
  r_rs_b_long <- get(data("r_rs_b_gamache", package = pkgname), envir = environment())

  luts <- build_saber_stan_luts(
    wavelength = wavelength,
    a_w_df = a_w_df,
    a0_a1_df = a0_a1_df,
    r_rs_b_long = r_rs_b_long
  )

  bottom_class_ids <- match(bottom_class_names, luts$classes)
  if (any(is.na(bottom_class_ids))) {
    missing <- bottom_class_names[is.na(bottom_class_ids)]
    stop("Bottom class not found in library: ", paste(missing, collapse = ", "))
  }

  list(
    n_wl = length(wavelength),
    wavelength = wavelength,
    rrs_obs = rrs_obs,
    sigma = sigma,

    a_w = luts$a_w,
    a0 = luts$a0,
    a1 = luts$a1,
    bb_w = luts$bb_w,

    n_class = length(luts$classes),
    r_rs_b_lib = luts$r_rs_b_lib,
    bottom_class_ids = bottom_class_ids,

    water_type = water_type,
    theta_sun_deg = theta_sun_deg,
    theta_view_deg = theta_view_deg,
    shallow = shallow
  )
}


stan_data <- make_saber_stan_data(
  rrs_obs = rrs$rrs_0m,
  wavelength = rrs$wavelength,
  sigma = rrs$rrs_unc,
  water_type = 2,
  theta_sun_deg = 20,
  theta_view_deg = 0,
  shallow = 1,
  bottom_class_names = c("saccharina", "cca", "gravel"),
  pkgname = "SABER")
```

```{r}
# bottom_class_id = c(1 = "r_rs_b_saccharina", 2 = "r_rs_b_cca", 3 = "r_rs_b_gravel")

fit <- model$sample(
  data = stan_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 500,
  iter_sampling = 500
)
```

Now we can compare what those OAC gives in "am03" and compare that with our input rrs.

```{r}
library(posterior)
library(plotly)
library(latex2exp)
library(glue)

# --- Extract posterior draws for rrs_hat ---
draws_hat <- fit$draws(variables = "rrs_hat")
rrs_hat_mat <- as_draws_matrix(draws_hat)  # draws x n_wl columns

rrs_hat_mean <- colMeans(rrs_hat_mat)
rrs_hat_q <- apply(rrs_hat_mat, 2, quantile, probs = c(0.05, 0.95))
rrs_hat_lo <- rrs_hat_q[1, ]
rrs_hat_hi <- rrs_hat_q[2, ]

# --- Parameter summary for annotation (posterior mean) ---
# Added: a_g_s, a_nap_s, bb_p_gamma
par_vars <- c(
  "chl", "a_g_440", "a_g_s",
  "a_nap_440", "a_nap_s",
  "bb_p_550", "bb_p_gamma",
  "h_w", "bottom_frac"
)

draws_par <- fit$draws(variables = par_vars)
par_mean <- colMeans(as_draws_matrix(draws_par))

# keep it readable (optional)
sub_par <- par_mean
sub_par <- sub_par[order(names(sub_par))]

annot_text <- paste(glue("{names(sub_par)}: {round(sub_par, 3)}"), collapse = "\n")

# --- Plotly ---
ply <- plot_ly() %>%
  # Observed
  add_lines(x = rrs$wavelength, y = rrs$rrs_0m, name = "rrs_observed") %>%

  # 90% band (filled polygon)
  add_ribbons(
    x = rrs$wavelength,
    ymin = rrs_hat_lo,
    ymax = rrs_hat_hi,
    name = "rrs_hat 90% CI",
    hoverinfo = "skip",
    line = list(width = 0)
  ) %>%

  # Posterior mean forward model
  add_lines(
    x = rrs$wavelength, y = rrs_hat_mean,
    name = "rrs_hat mean",
    line = list(width = 3)
  ) %>%

  # Annotation block
  add_annotations(
    text = annot_text,
    textangle = 0,
    x = 1,
    xshift = 0,
    y = 1,
    yref = "paper",
    xref = "paper",
    xanchor = "top-left",
    yanchor = "top-left",
    showarrow = FALSE,
    font = font
  ) %>%

  layout(
    yaxis = list(title = TeX("R_\\text{rs}(0^-) [\\text{sr}^{-1}]")),
    xaxis = list(title = TeX("\\text{wavelength} [\\text{nm}]")),
    font = font,
    shapes = list(border),
    legend = list(
      font = list(size = 20),
      bgcolor = "rgba(105, 159, 245, 0.0)",
      xanchor = "top-right",
      yanchor = "top-right",
      x = 0.1,
      y = 1
    )
  ) %>%
  config(mathjax = "cdn", displayModeBar = TRUE)

ply
```

### Bottom reflectance retrieval

```{r}
r_rs_b_retrieved_unconstrained <- retrieve_r_rs_b_am03(
  wavelength = rrs$wavelength,
  iop = iop_unconstrained,
  r_rs = rrs$rrs_0m,
  water_type = 2,
  theta_view = 0,
  theta_sun = 20,
  h_w = par_unconstrained["h_w"]
)

ply <- plot_ly() %>%
  add_lines(x = rrs$wavelength, y = r_rs_b_retrieved_unconstrained, name = "r_rs_b_unconstrained") %>%
  layout(
    yaxis = list(
      title = TeX("R_\\text{rs}(\\text{bottom}) [sr^{-1}]"),
      range = c(-0.01, 0.6)
      ),
    xaxis = list(
      title = TeX("\\text{wavelength}[\\text{nm}]")
      ),
    font = font,
    shapes = list(border)
  ) %>%
  config(mathjax = "cdn", displayModeBar = T)

ply
# save_image(ply, here("2_chapter/figure/retrieved_bottom_reflectance_unconstrained.png"), width = 800, height = 700, scale = 4)
```

### Batch processing

```{r}
library(posterior)

# Fast-ish quantiles (posterior::quantile2 avoids some overhead)
qfun <- function(x, probs) posterior::quantile2(x, probs = probs)

summarize_rrs_hat <- function(fit, probs = c(0.05, 0.95), chunk_wl = 64) {
  # Pull only rrs_hat, keep as matrix: draws x (n_obs*n_wl)
  mat <- as_draws_matrix(fit$draws(variables = "rrs_hat"))
  cn  <- colnames(mat)

  # Parse indices rrs_hat[n,wl]
  idx <- do.call(
    rbind,
    strcapture("rrs_hat\\[(\\d+),(\\d+)\\]", cn,
               proto = list(n = integer(), wl = integer()))
  )
  n_obs <- max(idx[, "n"])
  n_wl  <- max(idx[, "wl"])

  # Order columns by (n, wl) so we can reshape cheaply
  o <- order(idx[, "n"], idx[, "wl"])
  mat <- mat[, o, drop = FALSE]
  idx <- idx[o, , drop = FALSE]

  # Preallocate outputs
  mean_hat <- matrix(NA_real_, n_obs, n_wl)
  qs_hat   <- array(NA_real_, dim = c(n_obs, n_wl, length(probs)),
                    dimnames = list(NULL, NULL, paste0("q", probs*100)))

  # Process by wavelength chunks (better cache/memory behavior)
  # columns are ordered by n then wl, so each wl forms a regular stride
  for (wl0 in seq(1, n_wl, by = chunk_wl)) {
    wl1 <- min(n_wl, wl0 + chunk_wl - 1)

    cols <- which(idx[, "wl"] >= wl0 & idx[, "wl"] <= wl1)
    sub  <- mat[, cols, drop = FALSE]

    m <- colMeans(sub)
    q <- apply(sub, 2, qfun, probs = probs)  # returns length(probs) x ncol

    # Reshape back: because ordered by (n,wl), we can fill by index quickly
    n_idx  <- idx[cols, "n"]
    wl_idx <- idx[cols, "wl"]
    mean_hat[cbind(n_idx, wl_idx)] <- m
    for (k in seq_along(probs)) {
      qs_hat[cbind(n_idx, wl_idx, k)] <- q[k, ]
    }
  }

  list(mean = mean_hat, quantiles = qs_hat)
}

summarize_params <- function(fit, par_vars,
                             probs = c(0.05, 0.5, 0.95)) {
  summarise_draws(
    fit$draws(variables = par_vars),
    mean, sd,
    ~posterior::quantile2(.x, probs = probs),
    rhat, ess_bulk, ess_tail
  )
}

fit_metrics <- function(rrs_obs, mean_hat, q05, q95) {
  stopifnot(all(dim(rrs_obs) == dim(mean_hat)))

  resid <- rrs_obs - mean_hat
  rmse <- sqrt(rowMeans(resid^2))

  # Spectral Angle Mapper (SAM)
  dotp <- rowSums(rrs_obs * mean_hat)
  n1   <- sqrt(rowSums(rrs_obs^2))
  n2   <- sqrt(rowSums(mean_hat^2))
  sam  <- acos(pmax(-1, pmin(1, dotp / (n1 * n2))))

  # Coverage of 90% CI (or whatever probs you used)
  coverage <- rowMeans(rrs_obs >= q05 & rrs_obs <= q95)

  data.frame(rmse = rmse, sam = sam, coverage = coverage)
}

postprocess_stan <- function(fit, rrs_obs, par_vars,
                             spec_probs = c(0.05, 0.95),
                             chunk_wl = 64) {

  spec <- summarize_rrs_hat(fit, probs = spec_probs, chunk_wl = chunk_wl)

  mean_hat <- spec$mean
  q_lo <- spec$quantiles[, , 1]
  q_hi <- spec$quantiles[, , 2]

  metrics <- fit_metrics(rrs_obs, mean_hat, q_lo, q_hi)
  par_sum <- summarize_params(fit, par_vars)

  list(
    spectrum_mean = mean_hat,
    spectrum_q_lo = q_lo,
    spectrum_q_hi = q_hi,
    metrics = metrics,
    param_summary = par_sum
  )
}


```


```{r}
ensembles <- unique(rrs_hw$ensemble)

ensembles <- ensembles[
  ensembles > as.numeric(ymd_hms("2022-07-06 20:58:20")) &
    ensembles > as.numeric(ymd_hms("2022-07-06 21:04:50"))
]

rrs_hw_sub <- rrs_hw %>%
  filter(wavelength > 400, wavelength < 720) %>%
  mutate(
    rrs_0m = rrs_0p_to_0m(rrs_estimate),
    h_w_meas = altitude_m + abs(bottom_elevation_m)
  )

rrs_hw_sub <- rrs_hw_sub %>%
  group_by(ensemble) %>%
  nest() %>%
  mutate(
    pass = purrr::map_lgl(data, ~ any(.x$rrs_unc > 0.001, na.rm = TRUE))
  ) %>%
  filter(pass == FALSE)# & ensemble %in% ensembles)
```

```{r}
run_one_ensemble <- function(df, ensemble_id, mod, stan_data_base,
                             out_dir,
                             iter_warmup = 500, iter_sampling = 500,
                             chains = 4, seed = 123,
                             use_measured_sigma = TRUE,
                             spec_probs = c(0.05, 0.95),
                             keep_fit = FALSE) {

  stopifnot(dir.exists(out_dir))

  # --- Inputs aligned to Stan wavelength grid ---
  rrs_obs <- as.numeric(df$rrs_0m)
  wl      <- as.numeric(df$wavelength)

  stopifnot(length(wl) == stan_data_base$n_wl)
  stopifnot(all(abs(wl - stan_data_base$wavelength) < 1e-8))


  
  sigma_vec <- if (use_measured_sigma) {
    as.numeric(df$rrs_unc)          # already 1 SD per wavelength
  } else {
    rep(0.005, stan_data_base$n_wl)  # fallback
  }
  
  # safety floor
  sigma_vec[!is.finite(sigma_vec) | sigma_vec <= 0] <- NA_real_
  sigma_floor <- 1e-12
  sigma_vec <- pmax(sigma_vec, sigma_floor)
  
  stopifnot(length(sigma_vec) == stan_data_base$n_wl)
  
  stan_data <- within(stan_data_base, {
    rrs_obs <- rrs_obs
    sigma   <- sigma_vec
  })

  # --- Output folder per ensemble ---
  ens_dir <- file.path(out_dir, paste0("ensemble_", ensemble_id))
  dir.create(ens_dir, recursive = TRUE, showWarnings = FALSE)

  # --- Sample, saving CSVs into the ensemble folder (best way to "save Stan result") ---
  fit <- mod$sample(
    data = stan_data,
    seed = seed,
    chains = chains,
    parallel_chains = chains,
    iter_warmup = iter_warmup,
    iter_sampling = iter_sampling,
    refresh = 0,
    output_dir = ens_dir
  )

  # Save minimal provenance
  # saveRDS(
  #   list(ensemble_id = ensemble_id,
  #        wl = wl,
  #        sigma = sigma,
  #        iter_warmup = iter_warmup,
  #        iter_sampling = iter_sampling,
  #        chains = chains,
  #        seed = seed),
  #   file = file.path(ens_dir, "run_info.rds")
  # )
  # saveRDS(stan_data, file = file.path(ens_dir, "stan_data.rds"))

  # --- Diagnostics (fast, robust) ---
  summ_all <- fit$summary()

  diag_draws <- fit$sampler_diagnostics()
  diag_mat   <- posterior::as_draws_matrix(diag_draws)

  n_divergent <- if ("divergent__" %in% colnames(diag_mat)) sum(diag_mat[, "divergent__"]) else NA_integer_

  diag <- list(
    n_divergent = n_divergent,
    max_rhat = max(summ_all$rhat, na.rm = TRUE),
    min_ess_bulk = min(summ_all$ess_bulk, na.rm = TRUE)
  )
  saveRDS(diag, file = file.path(ens_dir, "diagnostics.rds"))

  # --- Parameter posterior summaries (mean/sd/quantiles + rhat/ess) ---
  par_vars <- c("chl", "a_g_440", "a_nap_440", "bb_p_550", "h_w", "bottom_frac")

  par_sum <- posterior::summarise_draws(
    fit$draws(variables = par_vars),
    mean, sd,
    ~posterior::quantile2(.x, probs = c(0.05, 0.5, 0.95)),
    rhat, ess_bulk, ess_tail
  )
  saveRDS(par_sum, file = file.path(ens_dir, "param_summary.rds"))

  # A wide "means" row (handy for quick joins)
  par_mean_wide <- try({
    par_sum |>
      dplyr::filter(.data$variable %in% c("chl","a_g_440","a_nap_440","bb_p_550","h_w") |
                      grepl("^bottom_frac\\[", .data$variable)) |>
      dplyr::select(.data$variable, .data$mean) |>
      tidyr::pivot_wider(names_from = .data$variable, values_from = .data$mean)
  }, silent = TRUE)
  if (!inherits(par_mean_wide, "try-error")) {
    saveRDS(par_mean_wide, file = file.path(ens_dir, "param_mean_wide.rds"))
  } else {
    par_mean_wide <- NULL
  }

  # --- Spectrum post-processing: mean + quantiles + fit metrics ---
  # Works for rrs_hat[wl] OR rrs_hat[1,wl]
  rrs_out <- NULL
  draws_hat <- try(fit$draws(variables = "rrs_hat"), silent = TRUE)

  if (!inherits(draws_hat, "try-error")) {
    mat <- posterior::as_draws_matrix(draws_hat)
    cn  <- colnames(mat)

    # Parse either [wl] or [n,wl]
    idx1 <- suppressWarnings(
      strcapture("rrs_hat\\[(\\d+)\\]$", cn, proto = list(wl = integer()))
    )
    idx2 <- suppressWarnings(
      strcapture("rrs_hat\\[(\\d+),(\\d+)\\]$", cn, proto = list(n = integer(), wl = integer()))
    )

    if (!all(is.na(idx1[, "wl"]))) {
      wl_idx <- idx1[, "wl"]
    } else if (!all(is.na(idx2[, "wl"]))) {
      # take n==1 if present; otherwise just use wl
      keep <- is.na(idx2[, "n"]) | idx2[, "n"] == 1
      mat  <- mat[, keep, drop = FALSE]
      wl_idx <- idx2[keep, "wl"]
      cn <- colnames(mat)
    } else {
      stop("Could not parse rrs_hat indices. Expected rrs_hat[wl] or rrs_hat[n,wl].")
    }

    o <- order(wl_idx)
    mat <- mat[, o, drop = FALSE]
    wl_idx <- wl_idx[o]

    # Summaries per wavelength
    rrs_hat_mean <- colMeans(mat)
    rrs_hat_q    <- apply(mat, 2, posterior::quantile2, probs = spec_probs)
    # rrs_hat_q: length(probs) x n_wl

    # Ensure full length n_wl (in case some indices missing)
    mean_full <- rep(NA_real_, stan_data_base$n_wl)
    q_lo_full <- rep(NA_real_, stan_data_base$n_wl)
    q_hi_full <- rep(NA_real_, stan_data_base$n_wl)

    mean_full[wl_idx] <- rrs_hat_mean
    q_lo_full[wl_idx] <- rrs_hat_q[1, ]
    q_hi_full[wl_idx] <- rrs_hat_q[2, ]

    # Fit metrics
    resid <- rrs_obs - mean_full
    rmse  <- sqrt(mean(resid^2, na.rm = TRUE))

    dotp <- sum(rrs_obs * mean_full, na.rm = TRUE)
    n1   <- sqrt(sum(rrs_obs^2, na.rm = TRUE))
    n2   <- sqrt(sum(mean_full^2, na.rm = TRUE))
    sam  <- acos(max(-1, min(1, dotp / (n1 * n2))))

    coverage <- mean(rrs_obs >= q_lo_full & rrs_obs <= q_hi_full, na.rm = TRUE)

    rrs_out <- list(
      wl = wl,
      mean = mean_full,
      q_lo = q_lo_full,
      q_hi = q_hi_full,
      probs = spec_probs,
      metrics = data.frame(rmse = rmse, sam = sam, coverage = coverage)
    )

    saveRDS(rrs_out, file = file.path(ens_dir, "rrs_posterior_summaries.rds"))
    saveRDS(rrs_out$metrics, file = file.path(ens_dir, "rrs_fit_metrics.rds"))
  }

  # Save a lightweight overall summary table (optional but handy)
  saveRDS(summ_all, file = file.path(ens_dir, "stan_summary_all.rds"))

  # If you want to persist the CmdStanMCMC object itself:
  # - the CSVs in ens_dir are already the canonical saved result
  # - you can also serialize the fit object, but it can be large
  if (keep_fit) {
    saveRDS(fit, file = file.path(ens_dir, "fit_object.rds"))
  }

  list(
    ensemble = ensemble_id,
    out_dir = ens_dir,
    diag = diag,
    par_mean = par_mean_wide,
    par_summary = par_sum,
    rrs = rrs_out,
    fit = if (keep_fit) fit else NULL
  )
}
```

```{r}
# run_one_ensemble(df, ensemble_id, mod, stan_data_base,
#                              out_dir,
#                              iter_warmup = 500, iter_sampling = 500,
#                              chains = 4, seed = 123,
#                              use_measured_sigma = TRUE,
#                              spec_probs = c(0.05, 0.95),
#                              keep_fit = FALSE)

results <- rrs_hw_sub %>%
  mutate(
    stan = purrr::map2(
      data,
      ensemble,
      ~ run_one_ensemble(
        .x, .y, model, stan_data, out_dir = "/D/Documents/phd/thesis/4_chapter/data/stan/"
          )
      )
  )
```

```{r}
par_inversed <- results %>%
  transmute(ensemble, par_mean = purrr::map(stan, "par_mean")) %>%
  tidyr::unnest(par_mean) %>%
  rename_with(~ paste0(.x, "_estim"), .cols = -ensemble)

diag_tbl <- results %>%
  transmute(
    ensemble,
    n_divergent = purrr::map_int(stan, ~ .x$diag$n_divergent),
    max_rhat = purrr::map_dbl(stan, ~ .x$diag$max_rhat),
    min_ess_bulk = purrr::map_dbl(stan, ~ .x$diag$min_ess_bulk)
  )
```


```{r}

compare_h_w <- rrs_hw_sub %>%
  unnest(cols = c(data)) %>%
  select(ensemble, lat, lon, h_w_meas, plant_height_m, percent_coverage) %>%
  distinct() %>%
  inner_join(par_inversed, by = "ensemble")

ply <- plot_ly(compare_h_w) %>%
  add_markers(
    x = ~h_w_meas,
    y = ~`h_w_estim`,
    # color = ~`a_g_440_estim`,
    text = ~ensemble
  ) %>%
  layout(
    font = font,
    xaxis = list(title = list(text = TeX("h_\\text{w}[\\text{m}]\\ \\text{measured}"))),
    yaxis = list(title = list(text = TeX("\\hat{h_\\text{w}}[\\text{m}]\\ \\text{retrieved}"))),
    shapes = list(
      border,
      list(
        type = "line",
        x0 = min(compare_h_w$h_w_meas, compare_h_w$`h_w_estim`, na.rm = TRUE),
        x1 = max(compare_h_w$h_w_meas, compare_h_w$`h_w_estim`, na.rm = TRUE),
        xref = "x",
        y0 = min(compare_h_w$h_w_meas, compare_h_w$`h_w_estim`, na.rm = TRUE),
        y1 = max(compare_h_w$h_w_meas, compare_h_w$`h_w_estim`, na.rm = TRUE),
        yref = "y",
        line = list(color = "black")
      )
    )
  ) %>%
  config(mathjax = "cdn", displayModeBar = TRUE)

ply
```

## Water depth constrained inversion

```{r}
# How to retrieve bottom reflectance without specifying a class ?
# Is it feasible ? Definitely seems useful.
par_fixed <- c(
  "water_type" = 2,
  "theta_view" = 0,
  "theta_sun" = 20,
  # "h_w" = 4.34,
  "h_w" = 4.44
)

par_constrained <- inverse_mcmc(
  rrs = rrs,
  forward_model = "am03",
          par_inversed = c("chl", "a_g_440", "a_g_s", "bb_p_550", "bb_p_gamma", "r_rs_b_saccharina", "r_rs_b_cca", "r_rs_b_gravel", "sd"),
          prior = NULL,
          lower = c(0, 0, 0, 0, 0, 0, 0, 0, 0.001),
          best = c(1, 4, 0.017, 0.04, 0.46, 0.5, 0.25, 0.25, 0.1),
          upper = c(20, 10, 1, 0.5, 5, 1, 1, 1, 1),
  par_fixed = par_fixed,
  iterations = 5e4,
  burnin = 2000,
  sampler = "DEzs"
)

# plot(out)
```

```{r}
# par_estimates <- tibble(
#   name = names(MAP(out)[[1]]),
#   value = MAP(out)[[1]]
# ) %>%
#   filter(name %in% c("chl", "a_g_440", "bb_p_550", "h_w"))
#
# par_estimates <- stats::setNames(par_estimates$value, par_estimates$name)

iop <- iop_from_oac(
  rrs$wavelength,
  par_constrained
)

r_b_fraction_vec <- par_constrained[grepl("^r_rs_b(?!.*_sd$)", names(par_constrained), perl = TRUE)]

rrs_bottom <- compute_r_rs_b_lmm(
  fractions = r_b_fraction_vec
)

rrs_constrained <- forward_am03(
  wavelength = rrs$wavelength,
  iop = iop,
  water_type = 2,
  theta_view = 0,
  theta_sun = 20,
  h_w = par_fixed["h_w"],
  r_b = rrs_bottom
)

sub_par <- par_constrained[1:8]

ply <- plot_ly() %>%
  add_lines(x = rrs$wavelength, y = rrs$rrs_0m, name = "rrs_observed") %>%
  add_lines(x = rrs$wavelength, y = rrs_constrained, name = "rrs_retrieved") %>% 
  add_annotations(
    text = ~paste(glue::glue("{names(sub_par)}: {round(sub_par, 3)}"), collapse = "\n"),
    textangle = 0,
    x = 1,
    xshift = 0,
    y = 1,
    yref = "paper",
    xref = "paper",
    xanchor = "top-left",
    yanchor = "top-left",
    showarrow = FALSE,
    font = font
  ) %>%
  layout(
    yaxis = list(
      title = TeX("R_\\text{rs}(0^-) [\\text{sr}^{-1}]")
    ),
    xaxis = list(
      title = TeX("\\text{wavelength} [\\text{nm}]")
    ),
    font = font,
    shapes = list(border),
    legend = list(
      font = list(
        size = 20),
      bgcolor = "rgba(105, 159, 245, 0.0)",#"#FFFFFF",
      xanchor = "top-right",
      yanchor = "top-right",
      x = 0.1,
      y = 1
    )
  ) %>%
  config(mathjax = "cdn", displayModeBar = T)

ply

# save_image(ply, here("2_chapter/figure/rrs_input_estimates_output_constrained_hw.png"), width = 800, height = 700, scale = 4)
```

```{r}
r_rs_b_retrieved_constrained <- retrieve_r_rs_b_am03(
  wavelength = rrs$wavelength,
  iop = iop,
  r_rs = rrs$rrs_0m,
  water_type = 2,
  theta_view = 0,
  theta_sun = 20,
  h_w = par_fixed["h_w"]
)

ply <- plot_ly() %>%
  add_lines(x = rrs$wavelength, y = r_rs_b_retrieved_constrained, name = "r_rs_b_constrained") %>%
  add_lines(x = rrs$wavelength, y = r_rs_b_retrieved_unconstrained, name = "r_rs_b_unconstrained") %>%
  add_lines(
    x = unique(r_rs_b_gamache$wavelength),
    y = r_rs_b_gamache$r_rs_b_mean[r_rs_b_gamache$class == "saccharina"],
    name = "saccharina") %>%
  layout(
    yaxis = list(
      title = TeX("R_\\text{rs}(\\text{bottom}) [sr^{-1}]"),
      range = c(-0.01, 0.06)
      ),
    xaxis = list(
      title = TeX("\\text{wavelength}[\\text{nm}]"),
      range = c(450, 720)
      ),
    legend = list(
      font = list(
        size = 20),
      bgcolor = "rgba(105, 159, 245, 0.0)",#"#FFFFFF",
      xanchor = "top-right",
      yanchor = "top-right",
      x = 0.5,
      y = 1
    ),
    font = font,
    shapes = list(border)
  ) %>%
  config(mathjax = "cdn", displayModeBar = T)

ply
# save_image(ply, here("2_chapter/figure/retrieved_bottom_reflectance_constrained_hw.png"), width = 800, height = 700, scale = 4)
```
