---
title: "water_surface_reflectance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{water_surface_reflectance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SABER)
library(ggplot2)
```

# Gege (2012)

We can test the Gege (2012) model with data shipped with SABER.

```{r}
rrs_subsurface <- test_input_rrs$Rrs
wavelength <- test_input_rrs$Wavelength
```

We define the model input parameter

```{r}
theta_s = 30 #sun angle of sensor above water surfac
theta_v = 15 #viewing angle of sensor above water surface
# Water surface roughness [sr]
q_surf = pi 
# Atmospheric conditions
# Irradiance intensities [1/sr]
g_dd=0.02
g_dsr=1/pi
g_dsa=1/pi
# Intensities of light sources
f_dd= 1
f_ds= 1
# Angstrom exponent
alpha = 1.317
# Atmospheric pressure [mbar]
P = 1013.25 
# Relative Humidity
RH = 0.60
# Scale height for ozone [cm]
Hoz = 0.300 
# Scale height of the precipitate water in the atmosphere [cm]
WV= 2.500
```

We resample the input extraterrestrial irradiance and atmospheric absorption spectrum of WASI.

```{r}
# Extraterrestrial solar irradiance [mW/m^2 nm]
F0_res <- Hmisc::approxExtrap(
  x =F0$wavelength,
  y =F0$F0,
  xout = wavelength,
  method = "linear")$y

# Oxygen absorption [1/cm]
a_O2_res <- Hmisc::approxExtrap(
  x =a_O2$wavelength,
  y =a_O2$a_O2,
  xout = wavelength,
  method = "linear")$y

# Ozone absorption [1/cm]
a_O3_res <- Hmisc::approxExtrap(
  x = a_O3$wavelength,
  y = a_O3$a_O3,
  xout = wavelength,
  method = "linear")$y

# Water vapour absorption [1/cm]
a_WV_res <- Hmisc::approxExtrap(
  x =a_WV$wavelength,
  y =a_WV$a_WV,
  xout = wavelength,
  method = "linear")$y
```

We compute the angle of refraction and Fresnel reflectance

```{r}

theta_s <- theta_s*(pi/180)
theta_v <- theta_v*(pi/180)

theta_v_t <- refraction_law(theta_v, n1 = 1, n2 = 1.34) 
theta_s_t <- refraction_law(theta_s, n1 = 1, n2 = 1.34) 
fresnel_rho_sun = fresnel_reflectance(theta_s, theta_s_t)
```

We use the semi-analytically model to transmit $F0$ to direct $E_\text{dd}(0^{+})$ and diffuse$E_\text{ds}(0^{+})$ downwelling irradiance just above the surface.

```{r}
# Atmospheric path length -------------------------------------------------
M <- 1/(cos(theta_s)+(0.50572*((90 + 6.079975-theta_s)^(-1.253))))
# M corrected for non standard atmopsheric pressure
M1 <-  (M*P)/1013.25
# path length for ozone
Moz <-  1.0035/ (((cos(theta_s)^2)+0.007)^0.5)

# Air mass type
# TODO: type_case_water is defined in main.R as an integer outside the
#   function. How was this working with this scope ?
type_case_water <- 1
if (type_case_water == "1") { 
  AM <- 1
} else{
  AM <- 1
}

# single scattering albedo
omega_a <- ((-0.0032*AM) + 0.972)*exp(RH*3.06*(10^-4))

# aerosol scale height [km]
Ha <- 1 
# horizontal visibility [km]
V <- 15
# 
beta <- 3.91*(Ha/V)
tau_a <-  beta*((wavelength/550)^(-alpha))

# Atmospheric transmittance spectra ---------------------------------------
# Rayleigh scattering
Tr <-   exp(-M1/((115.6406*(wavelength^(4)))-(1.335*(wavelength^(2)))))
# Aerosol absorption
Taa <-  exp(-(1-omega_a)*tau_a*M)
# Aerosol scattering
Tas <-  exp(-omega_a*tau_a*M)
# Ozone absorption
Toz <-  exp(-a_O3_res*Hoz*Moz)
# Oxygen absorption
To <-   exp((-1.41*a_O2_res*M1)/((1+(118.3*a_O2_res*M1))^0.45))
# Water vapor absorption
Twv <-  exp((-0.2385*a_WV_res*WV*M)/((1+(20.07*a_WV_res*WV*M))^0.45))

# Aerosol forward scattering probability ----------------------------------
# asymmetry factor of the aerosol scattering phase function
if (alpha < 0) {
  assymetry_factor <- 0.82
} else if (alpha > 1.2) {
  assymetry_factor <- 0.65
} else {
  assymetry_factor <- -0.1417*alpha+0.82
}
# Empirical coefficient for the Aerosol forward scattering probability
B3 <- log(1 - assymetry_factor)
B2 <-  B3*(0.0783 +(B3*(-0.3824-0.5874*B3)))
B1 <-  B3*(1.459 +(B3*(0.1595+0.4129*B3)))
# Aerosol forward scattering probability
Fa <-  1-(0.5*exp((B1+(B2*cos(theta_s)))*cos(theta_s)))

# Downwelling irradiances -------------------------------------------------
# Direct 
Edd <-  F0_res*Tr*Taa*Tas*Toz*To*Twv*cos(theta_s)
# Diffuse Rayleigh
Edsr <- (1/2)*F0_res*(1-(Tr^(0.95)))*Taa*Tas*Toz*To*Twv*cos(theta_s)
# Diffuse aerosol
Edsa <-  F0_res*(Tr^(1/2))*Taa*(1-Tas)*Toz*To*Twv*cos(theta_s)*Fa
# Diffuse downwelling irradiance above the surface
Eds <-  Edsr+Edsa

# Downwelling irradiance just above the water surface
Ed_surface <-  (f_dd*Edd) + (f_ds*Eds)
```

We compute the WASI4 equivalent "sky radiance", direct and diffuse component, $L_\text{skyd}$ and $L_\text{skys}$.
*Note that those radiance are not parametrized in term of viewing angles and therefore cannot be used to simulate a sensor observation, see: Gege, P. (2015) ‘The Water Colour Simulator WASI’.*

```{r}
# WASI sky radiances ------------------------------------------------------
# Sky Radiance [mW/sr m^2 nm]
Ls_direct <-  (g_dd*Edd)

Ls_diff <- (g_dsr*Edsr) + (g_dsa*Edsa)

# Sun glint radiance
L_sun_glint <- fresnel_rho_sun*Ls_direct

# Sky glint radiance
L_sky_glint <- fresnel_rho_sun*Ls_diff
```

Plot the result.

```{r}
ggplot() +
  geom_line(aes(x = wavelength, y = L_sun_glint/Ed_surface, colour = "sun glint")) +
  geom_line(aes(x = wavelength, y = L_sky_glint/Ed_surface, colour = "sky glint")) +
  geom_line(aes(x = wavelength, y = rrs_subsurface, colour = "Rrs")) + 
  geom_line(aes(x = wavelength, y = rrs_subsurface+(L_sun_glint/Ed_surface)+(L_sky_glint/Ed_surface), colour = "Rrs total"))

```

```{r}
result <- SABER::gege_2012(
  wavelength <- test_input_rrs$Wavelength,
  theta_s = 30,
  g_dd=0.01,
  g_dsr=0,
  g_dsa=0,
  f_dd= 1,
  f_ds= 1,
  alpha = 1.317,
  P = 1013.25,
  RH = 0.60,
  Hoz = 0.300, 
  WV= 2.500
)

ggplot(result) +
  #geom_line(aes(x = wavelength, y = Ed_surface, colour = "Es")) +
  geom_line(aes(x = wavelength, y = L_sun_glint, colour = "Ls"))
```



# Vanhellemeont 2018

input parameters 

```{r}
wavelength <- test_input_rrs$Wavelength
tau_r <- 0.2
theta_s <-30
theta_v <- 15
delta_phi <- 0
```

computation

```{r}
theta_s <- theta_s*(pi/180)
theta_v <- theta_v*(pi/180)

theta_s_t <- refraction_law(theta_s, n1 = 1, n2 = 1.34)
theta_v_t <- refraction_law(theta_v, n1 = 1, n2 = 1.34)

fresnel_s <- fresnel_reflectance(theta_s, theta_s_t)
fresnel_v <- fresnel_reflectance(theta_v, theta_v_t)

cos_theta_plus <- cos(theta_s) * cos(theta_v) - sin(theta_s) * sin(theta_v) * cos(delta_phi)

# omega <- acos(sqrt(cos_2_omega))
# omega <- acos(cos_2_omega)/2

rayleigh_phase_omega <- 0.75*(1+cos_theta_plus^2)

rayleigh_phase = (fresnel_s + fresnel_v) * rayleigh_phase_omega

rho_sky <- tau_r*rayleigh_phase*(4*cos(theta_s)*cos(theta_v))^-1
```


# 
